### diag
diag(n)
diag(x,n)    x is the element, n is the dimension.
```{r}
# examples
diag(3)

x = c(1,5,9)
diag(x)

diag(x,5)

diag(2,4)

2*diag(4)
```

### matrix

matrix(element,row,column)

```{r}

matrix(1,4,4)
matrix(1,4,3)

x = c(1,5,9)
matrix(x,4,3)
matrix(x,4,3,byrow = TRUE)

matrix(1,5)
matrix(1,5,1)
matrix(1,ncol=5)
matrix(1,1,5)
```

### c , cbind
```{r}
x1=c(1,2,3,4)
x2=c(4,5,6,7)
x3=c(7,8,9,10)
cbind(x1,x2,x3)
rbind(x1,x2,x3)
```



### Matrix arithmetic and operations

+,-,*,/   elementwise

%*%       matrix multiplication       need to be comformable


```{r}
A = matrix(1:12,3)
B = matrix(2,3,4)
C = matrix(c(2,4,5,7,6,3,9,1),4)
A
B
A+B
A*B
A/B
exp(A)
```


crossprod(x,y) is equivalent to t(x) %*% y
```{r}
A1=matrix(1:15,ncol=3)
A1%*%B 
```


### Matrix Functions
 
 dim(x), nrow(x), ncol(x) obatain matrix dimensions
 
```{r}
dim(A)
nrow(A)
ncol(A)
```
 
 t(x)  transpose of x
 c(x) get a row with all elements stacked up

```{r}
c(A)
dim(t(A))
length(c(A))
matrix(c(A),3,4)
```
 
  solve(x) inverse of x
```{r}
D = matrix(c(2,5,0,8,1,4,2,6,7,8,9,3,1,5,7,8),4)
D
solve(D)
```
  

 chol(x) cholesky decomposition: if U <- chol(x), then t(U) %*% U reconstructs x
```{r}
E =  matrix(c(2,-1,0,-1,2,-1,0,-1,2),3)
U = chol(E)
U
t(U)%*%U
```
 
 判断是否正定
```{r}
library(matrixcalc)
is.positive.definite(E)
```
 
 eigen(x) eigendecomposition  you get the eigenvector and eigenvalues
 `e$vectors%*%diag(e$values,length(e$values))%*%t(e$vectors)` reconstructs `X`
```{r}
eigen(E)
```
 



### Rowwise and columnwise computation #

 apply(X,       Array,matrix or data frame
      MARGIN,   1:rows, 2:columns, c(1,2): rows and columns
      FUN ）    Function to be applied
```{r}
A
apply(A, MARGIN =1, FUN=sum)
apply(A, MARGIN =2, FUN=sum)
apply(A, MARGIN =c(1,2), FUN=exp)
setequal(apply(A,c(1,2),exp),exp(A))
```


 sweep(X,MARGIN,STATS,FUN) for each row&col evaluate FUN(X,STATS[i]) and replace the original row.
```{r}
A
sweep(A,1,c(1:3),"+")
sweep(A,2,c(1:4),"+")
```
 


* `colMeans`, `rowMeans`, `colSums`, `rowSums`: exactly as it sounds.
* `scale(X)`: centre the matrix so that its column means are 0 and its column mean-squared value is 1; further arguments can be used to do one or the other only.
```{r}
scale(A)
```








